Examining Whether Approximate Visualizations Impact Perception of Corrected Visualizations

### 2016-12-08 paper/paper.tex
typo
Paper updates

\subsection{Visualizing Uncertainty and Human Perception of Visualization}

@@

\begin{figure*}[!ht]
\centering
\subfloat[Approximate visualization based on 0.2\% of the data. Note that ``New Yersey'' is missing and that ``Florida'' is the state with the most outgoing flights.]{\includegraphics[width=0.3\textwidth]{states5009.png}%
\label{fig:approx}}
\hfil
\subfloat[Visualization of precise data.]{\includegraphics[width=0.3\textwidth]{states10.png}%
\label{fig:precise}}
\hfil
\subfloat[Precise with difference visualization. The orange ticks show the user the approximate data they have previously seen.]{\includegraphics[width=0.3\textwidth]{states5009diff.png}%
\label{fig:diff}}
\caption{Visualizations of approximate and precise data of the number of flight per state. (c) highlights the difference to the approximate data.}
\label{fig:types}
\end{figure*}

@@

We used one month of flight data to generate the visualizations.
The precise visualization of number of flights per state uses data where the airline was ``JetBlue'' and the visualization of the flight per airline uses data from ``New York State''.
The approximate visualizations use a random subset of the data for the precise visualizations with replacement.
We then computed the approximate values by dividing the count values by the fraction of rows used in the sample.

@@

@@

We randomized visualization order, approximate sample, and study condition for each participant.
As previously stated, the variation level of the approximate visualizations varied by visualization.

@@

Analysis of each study question will have four independent variables: visualization number (interval), type of data displayed (categorical), level of approximation (interval), and presentation technique (categorical).
We anticipate visualization number and type of data displayed will have no effect on our results, as they do not pertain to our research questions.
Should they have no effect, we will then discard these terms from our analysis and not report on them further.
Further, each participant will be treated as a random effect.

We are asking three types of questions in our study with different dependent variables:
\begin{enumerate}
\item Was there a difference between the precise and approximate visualizations (categorical dependent variable)
\item Which airline or state had the most or least flights (categorical dependent variable)
\item How many flights were on a particular airline or from a particular state (interval dependent variable)
\item Confidence in answers to the above questions (interval dependent variable)
\end{enumerate}

\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Example statistical analysis table for categorical dependent variables}
\label{table_example}
\centering
\begin{tabular}{|c||c|c|c|c|}
\hline
Study parameter & Est. & SE & Z & p\\
\hline
\hline
Presentation Technique & & & & \\
\hline
Approximate Sample Amount & & & & \\
\hline
Technique * Sample Amount & & & & \\
\hline
\end{tabular}
\end{table}

@@

@@

We had no way of ensuring that participants did not take screenshots in order to answer later questions about the precise visualizations.

>>>>

\subsection{Visualizing Uncertainty and Human Perception of Visualizations}

@@

\begin{figure*}[!ht]
\centering
\subfloat[Approximate visualization based on 0.2\% of the data. Note that ``New Jersey'' is missing and that ``Florida'' is the state with the most outgoing flights.]{\includegraphics[width=0.3\textwidth]{states5009.png}%
\label{fig:approx}}
\hfil
\subfloat[Visualization of precise data.]{\includegraphics[width=0.3\textwidth]{states10.png}%
\label{fig:precise}}
\hfil
\subfloat[Precise with difference visualization. The orange ticks show the user the approximate data they have previously seen.]{\includegraphics[width=0.3\textwidth]{states5009diff.png}%
\label{fig:diff}}
\caption{Visualizations of approximate and precise data of the number of flight per state. (c) highlights the difference to the approximate data.}
\label{fig:types}
\end{figure*}

@@

We used one month of flight data to generate the visualizations.
The precise visualization of number of flights per state uses data where the airline was ``JetBlue'' and the visualization of the flight per airline uses data from ``New York State''.
In the study, we randomized the airline and state to focus on.
The approximate visualizations use a random subset of the data for the precise visualizations with replacement.
We then computed the approximate values by dividing the count values by the fraction of rows used in the sample.

@@

\todo{Talk about our participants.}

@@

We randomized visualization order, approximate sample, focus state, focus airline, and study condition for each participant.
As previously stated, the variation level of the approximate visualizations varied by visualization.

\subsubsection{Dependent Variables}

We asked three types of questions about each precise visualization, bulding on Brehmer and Munzner's task taxonomy \cite{brehmer2013multi}.
Participants were asked to recall (1) how many flights were there on or out of the focus state or airline, (2) the difference in flights between the focus state or airline and another state or airline in the visualization, and (3) which states or airlines were in the visualization.
We asked question (2) three times for each precise visualization with three different comparison airlines.
This repetition, as well as the repetition between state and airline, gave us repeated measures for each participant.

Each dependent variable has a "correct" answer (e.g., the data presented in the precise visualization).
We normalize our outcome (e.g., the answer participants gave) by the by the value of the "correct" answer, e.g.,

\[\frac{correct\_answer - answer\_given}{correct\_answer}\]

We refer to this value as the \textbf{measured bias}.

The experimental conditions inform how much variation there was between the precise and the approximate visualizations.
For each question we ask the participants, we compute the difference between the "correct" answer (e.g., the data presented in the precise visualization) and the "approximate" answer (e.g., the data presented in the precise visualization).
Again, we normalize this value by the "correct" answer, e.g.,

\[\frac{correct\_answer - approximate\_answer}{correct\_answer}\]

We refer to this value as the \textbf{expected bias}.
We treat the expected bias as an independent variable in our analysis.
We hypothesized that the greater the expected bias, thr greater the measured bias would be.
Meaning, if an approximate visualization differed substantially from the precise visualization, we expected participant's answers to be close to correct (e.g., a measured bias of near 0), but skew in the direction of the approximate visualization.
We anticipated that the measured bias would be less than the expected bias, as participants had seen the updated, correct visualization.

For question (3), the answers were calculated via a Jaccard Index.

@@

We analyzed our data through linear regression using the \texttt{lm} package in R, analyzing the results with a factorial ANOVA (\texttt{anova} in R).
We hypothesized that multiple factors would impact our outcome metrics, so we used multiple regression in our analyses.
We treated each participant as a random effect.

As expected, visualization order, focus state and focus airline had no effect on our results for any of the tests we conducted.
For the comparison questions, the comparison state or airline also did not have an effect on our results.
None of these independent variables pertain to our research questions, so we discard the terms from our analysis and do not report on them further.

\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Regression table for Question 1. Baseline is only the precise visualization.}
\label{table_example}
\centering
\begin{tabular}{|c||c|c|c|c|}
\hline
Study parameter & Est. & SE & t & p\\
\hline
\hline
Expected Bias & & & & \\
\hline
Control condition & & & & \\
\hline
Condition:Difference & & & & \\
\hline
Condition:Both & & & & \\
\hline
Expected Bias * Condition:Difference & & & & \\
\hline
Expected Bias * Condition:Both & & & & \\
\hline
\end{tabular}
\end{table}

@@

\todo{future research questions}

@@

We had no way of ensuring that participants did not take screenshots in order to answer later questions about the precise visualizations.
\todo{some more detail here, now that we collected data}

\todo{Statistical methods and outliers}
